### 2014-11-13 intro.tex

The CLM language features two kinds of coordination primitives that can be used in the same
way as any other primitive, i.e., they are specified with the same
syntax and semantics as the rest of the programming language. These coordination
primitives can be used to improve program execution based on the state of the
program and the underlying machine. The first kind of coordination facts are
called \emph{sensing facts} and are used to sense information about the system
the program is running on, e.g., scheduling and node placement on threads. The
second kind of coordination facts are \emph{action facts} that when detected in
rules, are used to apply a scheduling operation during execution. Coordination
facts allow the programmer to write logical rules that depend on the current
state of the program and then prioritize node computation or place nodes in
different threads.

>>>>

The CLM language features two kinds of coordination primitives that can be used in the same
way as any other primitive, i.e., they are specified with the same
syntax and semantics as the rest of the programming language. These coordination
primitives can be used to improve program execution based on the state of the
program and the underlying machine. The first kind of coordination primitives are
called \emph{sensing facts} and are used to sense information about the system
the program is running on, e.g., scheduling and node placement on threads. The
second kind of coordination primitives are \emph{action facts} that when detected in
rules, are used to apply a scheduling operation during execution. Coordination
facts allow the programmer to write logical rules that depend on the current
state of the program and then prioritize node computation or place nodes in
different threads.

### 2014-11-13 related.tex

\paragraph{Declarative Languages}
Programming paradigms such as \emph{logic} and \emph{functional}
programming have been extensively exploited for their implicit parallelism.
In logic languages such as Prolog, researchers took advantage of the
non-determinism of proof-search to evaluate subgoals in parallel with models
such as \emph{or-parallelism}~\cite{ali-86} and
\emph{and-parallelism}~\cite{Shen-92}.  In functional languages, the stateless
nature of computation allows multiple expressions to evaluate safely in
parallel. This has been explored in several languages such as
NESL~\cite{Blelloch:1996:PPA:227234.227246}, Id~\cite{Nikhil93anoverview}, and
more recently Data Parallel Haskell~\cite{nepal}.
NESL often got good performance, but was limited to a particular
application space. We also limit our application space, but also give the
programmer a declarative way to coordinate computation to improve the
performance of the program.

>>>>

\paragraph{Declarative Languages}
Programming paradigms such as \emph{logic} and \emph{functional}
programming have been extensively exploited for their implicit parallelism.
In logic languages such as Prolog, researchers took advantage of the
non-determinism of proof-search to evaluate subgoals in parallel with models
such as \emph{or-parallelism}~\cite{ali-86} and
\emph{and-parallelism}~\cite{Shen-92}.  In functional languages, the stateless
computation allows multiple expressions to evaluate safely in
parallel. This has been explored in several languages such as
NESL~\cite{Blelloch:1996:PPA:227234.227246}, Id~\cite{Nikhil93anoverview}, and
more recently Data Parallel Haskell~\cite{nepal}.
NESL often got good performance, but was limited to a particular
application space. We also limit our application space, but also give the
programmer a declarative way to coordinate computation.
